Research Notes

### 2018-10-07 writeups/notes.tex
check in new meeting notes
stash

@@

One can solve \autoref{eq:shear_lin_uz} via \emph{Frobenius Theory}, but if we
consider only $\omega_{cr} \to 0$, we obtain a very simple power series
solution. In $\omega_{cr} \to 0$, we have only leading order contributions
\begin{equation}
    \ptd{u_{1z}}{z} + \frac{k_x^2N^2}{\omega_{cr}^2}u_{1z} = 0.
\end{equation}
Consider a power series solution $u_{1z} \propto \omega_{cr}^\alpha$, where
$\omega_{cr}(z) = \omega - k_xu_0(z)$. Note that very near the critical level,
$\pd{u_0}{z}\delta z k_x$ where $\delta z = z - z_c, \omega_{cr}(z_c) = 0$. We
thus have
\begin{align}
    \p*{\pd{\omega_{cr}}{z}}^2 \alpha\p*{\alpha - 1} + k_x^2N^2 &= 0,\\
    \alpha &= \frac{1}{2}\pm i\sqrt{\frac{N^2}{\p*{\pd{u_0}{z}}^2} -
        \frac{1}{4}}.
\end{align}
The classic Booker \& Bretherton paper notates this result $u_{1z} \propto
\p*{\omega_{cr}(z)}^{\frac{1}{2} \pm i\mu} \propto (\delta z)^{\frac{1}{2} \pm
i\mu}$.

>>>>

\section{Navier-Stokes Regularization}

This was a major tripping point for quite a while. As the wave breaks to higher
wavenumbers, it is expected that energy will accumulate in the higher $k$
spectral modes, and so some numerical viscosity must be added. This is not a
physical viscosity but a numerical one (since spectral codes have no numerical
viscosity of their own), and so the viscous term is added to all dynamical
equations $\rd{}{t} \to \pd{}{t} + \vec{u} \cdot \vec{\nabla} - \nu \nabla^2$.
This makes the equations resemble the Navier-Stokes equations, but since $\nu$
is not grounded in any physical process, it is not helpful to think of it as a
Navier-Stokes equation.

We choose the $\nu \nabla^2$ term, which scales $\sim k^2$, such that it becomes
strong in the same regime that nonlinearity $\p*{\vec{u} \cdot \vec{\nabla}}$
becomes important. Consider some $k_{turb}$ turbulent wavenumber, then we will
want $\nu k_{turb}^2 \sim uk_{turb}$. We can get an idea of what $u$ will be
when turbulence sets in using the wave breaking criterion, that breaking sets in
when $u \sim \omega/k$ where $k$ is the base wavelength of the wave. This
substitution suggests we should use $\nu = \frac{\omega}{k_{turb}k}$. We can
choose $k_{turb} = \frac{2\pi L}{N}$ the grid spacing, so that at this turbulent
wavelength the regularization kicks in. If we then take $k, \omega$ to be the
parent wave's properties, we have a prescription for $\nu$.

This turns out to be insufficient for regularization, we need roughly $\nu \sim
\frac{3\omega_0}{k_zk_{turb}}$.

This is because we performed the naive substitution $k, \omega$ from the parent
wave. Instead, it is advisable to scale $k, \omega$ to the same $k_{turb}$
length scale. We make ansatz that the cascade is roughly Kolmogorov, and so
(TODO work this out) we obtain that we should choose $\frac{\omega_0}{k_z^2 \nu}
\sim \p*{\frac{k_{turb}}{k_z}}^{4/3}$; this comes from scaling $\omega, k$ per
the Kolmogorov prescription $\omega_l u_l^2$ is approximately constant over
length scale $l$.

\subsection{Gotcha: $\nu_x, \nu_z$ and Grid Spacing}

If the grid resolution is uneven in the $x, z$ directions, it might be tempting
to use different $\nu_x, \nu_z$ per the above criteria, since their $k_{turb} =
k_{\max}$ differ. This is dangerous however, since $\nu \nabla^2 \vec{u}_1$ has
to be divergenceless! Thus, it is more advisable to choose an equal grid spacing
in the two directions.

@@

One can solve \autoref{eq:shear_lin_uz} via \emph{Frobenius Theory}, but if we
consider only $\omega_{cr} \to 0$, we obtain a very simple power series
solution. In $\omega_{cr} \to 0$, we have only leading order contributions
\begin{equation}
    \ptd{u_{1z}}{z} + \frac{k_x^2N^2}{\omega_{cr}^2}u_{1z} = 0.
\end{equation}
Consider a power series solution $u_{1z} \propto \omega_{cr}^\alpha$, where
$\omega_{cr}(z) = \omega - k_xu_0(z)$. Note that very near the critical level,
$\pd{u_0}{z}\delta z k_x = \omega_{cr}$ where $\delta z = z - z_c,
\omega_{cr}(z_c) = 0$. We thus have
\begin{align}
    \p*{\pd{\omega_{cr}}{z}}^2 \alpha\p*{\alpha - 1} + k_x^2N^2 &= 0,\\
    \alpha &= \frac{1}{2}\pm i\sqrt{\frac{N^2}{\p*{\pd{u_0}{z}}^2} -
        \frac{1}{4}}.
\end{align}
The classic Booker \& Bretherton paper notates this result $u_{1z} \propto
\p*{\omega_{cr}(z)}^{\frac{1}{2} \pm i\mu} \propto (\delta z)^{\frac{1}{2} \pm
i\mu}$.
